Bringing the Power of eBPF to Open vSwitch

### 2018-10-23 paper.tex
Edit ch 01

eBPF, extended Berkeley Packet Filter, enables userspace applications to
customize and extend the Linux kernel's functionality.  It provides
flexible platform abstractions for network functions, and is being ported
to a variety of platforms.  Among the various different ways of using eBPF,
OVS has been exploring the power of eBPF in three:
1) in-kernel flow processing by attaching eBPF programs to TC,
2) offloading a subset of processing to XDP (eXpress Data Path),
and 3) moving the flow processing to userspace by using AF\_XDP.

Attaching eBPF to TC, OVS-eBPF project, started first with the most aggressive goal:
we planned to re-implement the entire features of OVS kernel datapath under net/openvswitch/* into eBPF code.
We worked around a couple of limitations, for example, the lack of TLV (Type-Length-Value) support
led us to redefine a binary kernel-user API using a fixed-length array;
and without a dedicated way to execute a packet, we created a dedicated
device for user to kernel packet transmission,
with a different BPF program attached to handle packet execute logic.
Currently, OVS-eBPF can satisfy most of the basic feature requirement
for flow processing, including tunneling protocol support.
But more complicated processing such as connection tracking, NAT,
(de)fragmentation, and ALG are still under discussion.

Moving one layer below TC is called XDP (eXpress Data Path), a much faster layer for packet
processing, but with almost no extra packet metadata and limited BPF helpers support.
Depending on the complexity of flows, OVS can offload a subset of its flow processing to XDP
when feasible. However, the fact that XDP has fewer helper function support implies that either
1) only very limited number of flows are eligible for offload, or 2) more flow processing logic
needed to be done in native eBPF code.

XDP provides another way for interacting with userspace programs, called AF\_XDP. 
AF\_XDP is another socket interface for control plane and a shared
memory API for accessing packets from userspace application.
OVS today has another full-fledged
datapath implementation in userspace, called dpif-netdev, mostly used by DPDK community.
OVS-AFXDP project treats the AF\_XDP as a fast packet-I/O channel, and move
most of the flow processing into the userspace OVS.

A rough comparison of the two project is shown in Table~\ref{compare}
This paper describes the design, implementation, and evaluation of the
OVS-eBPF and OVS-AFXDP projects.
\begin{table}
\centering
\scriptsize
\begin{tabular}{|c | c | c|}
\hline
 {\bf Projects} & {\bf OVS-eBPF } & {\bf OVS-AFXDP} \\ \hline\hline
  Packet Rate & Low & High \\ \hline
  Driver Support & No & Yes \\ \hline
  Minimal Kernel Req. & 4.18 & Unknown \\ \hline
  BPF Code Size & Large & Minimal \\ \hline
  Extensibility & Low & High \\ \hline
  \end{tabular}
\caption{\footnotesize
Comparison of the OVS-eBPF project and OVS-AFXDP project.
}
\label{compare}
\end{table}

\section{Background}

@@

\section{OVS eBPF Datapath}
\myfigcross{ovsebpf}{5.5in}{The overall architecture of OVS eBPF datapath
consists of multiple eBPF programs which are tail-called dynamically, maps
which are shared between eBPF programs and userspace applications, and
ovs-vswitchd as the management plane for all components.}

@@

\section{Userspace OVS with AF\_XDP}
\begin{figure}
{\scriptsize
\begin{verbatim}
              |   +-------------------+
              |   |    ovs-vswitchd   |<-->ovsdb-server
              |   +-------------------+
              |   |      ofproto      |<-->OpenFlow controllers
              |   +--------+-+--------+ 
              |   | netdev | |ofproto-|
    userspace |   +--------+ |  dpif  |
              |   | AF_XDP | +--------+
              |   | netdev | |  dpif  |
              |   +---||---+ +--------+
              |       ||     |  dpif- |
              |       ||     | netdev |
              |_      ||     +--------+  
                      ||         
               _  +---||-----+--------+
              |   | XDP program +     |
 Linux kernel |   |   eBPF map        |
              |_  +--------||---------+
                           ||
                        physical
                           NIC
\end{verbatim}
}
\vspace{-1.0em}
\caption{OVS Architecture AF\_XDP}
\label{ovsafxdp}
\vspace{-1.0em}
\end{figure}

@@

\section{Conclusion}
This paper describes two eBPF projects related to OVS: OVS-eBPF and OVS-AFXDP.

>>>>

eBPF, extended Berkeley Packet Filter, enables userspace applications to
customize and extend the Linux kernel's functionality. It provides
flexible platform abstractions for network functions, and is being ported
to a variety of platforms. In Linux kernel, users can attach eBPF programs
to TC and XDP hook point as shown in Fig~\ref{ebpf-hook}. Based on this
design, we explore the possibilities of utilizing eBPF to implement OVS
datapath in threefold:
1) in-kernel flow processing by attaching eBPF programs to TC,
2) offloading a subset of flow processing to XDP (eXpress Data Path),
and 3) moving the flow processing to userspace by using AF\_XDP.

+-----------------------+
|    Network Stack      |
+---------^-------------+
          |
+---------|-------------+
| Traffic Control (TC)  | --> eBPF TC hook
+---------^-------------+
          |
+---------|-------------+
|     Driver (XDP)      | --> eBPF XDP hook
+---------^-------------+
          |
+---------|-------------+
|     Network card      |
+---------^-------------+
          |
          | packets

\end{verbatim}
}
\vspace{-1.0em}
\caption{eBPF program hook points in the Linux kernel.}
\label{ebpf-hook}
\vspace{-1.0em}
\end{figure}

Firstly, in OVS-eBPF project, we attach flow processing eBPF programs to TC.
We start with the most aggressive goal that we plan to re-implement the entire
features of OVS kernel datapath under net/openvswitch/* into eBPF code.
We work around a couple of eBPF limitations, for example, the lack of TLV
(Type-Length-Value) support leads us to redefine a binary kernel-user API using
a fixed-length array; without a dedicated way to execute a packet, we
create a dedicated device that attached a eBPF program to handle packet
execute logic for userspace to kernel packet transmission.
Currently, OVS-eBPF can satisfy most of the basic feature for flow processing
and tunneling protocol support, and we are investigating more complicated
features such as connection tracking, NAT, (de)fragmentation, and ALG.

We can attach eBPF programs one layer below TC to XDP (eXpress Data Path).
It is a much faster layer for packet processing, but there is almost no extra
packet metadata available, and XDP only provides limited kernel helpers.
Depending on the type of flows, OVS can offload a subset of its flow processing
to XDP. However, the fact that XDP has fewer helper function support implies
that either 1) only very limited number of flows are eligible for offload,
or 2) more flow processing logic needed to be implemented in native eBPF code.
For example, it is more difficult for OVS datapath to provide tunnel support in
XDP, since lightweight tunnel kernel helpers is not available.

XDP provides another way for interacting with userspace programs, called AF\_XDP. 
AF\_XDP is a socket interface for control plane and a shared memory API for
accessing packets from userspace application. With AF\_XDP, OVS-AFXDP project
redirect packets to userspace, and processes the packets by OVS's full-fledged
userspace datapath implementation, dpif-netdev. In this approach, OVS-AFXDP
project treats the AF\_XDP as a fast packet-I/O channel, and move most of
the flow processing into the userspace OVS.

This paper focuses on the OVS-eBPF and OVS-AFXDP projects.
A rough comparison of the two project is shown in Table~\ref{proj-compare}.
The remainder of this paper is organized as follows:
In Section~\ref{sec:background}, we first provide some background information
on eBPF, XDP, and AF\_XDP.  We then present the design, implementation, and
evaluation of the OVS-eBPF and OVS-AFXDP projects in Section~\ref{sec:ebpf}
and Section~\ref{sec:af_xdp} respectively. Finally,
Section~\ref{sec:conclusion} concludes the paper and discusses the future work.

\begin{table}
\centering
\scriptsize
\begin{tabular}{|c | c | c|}
\hline
 {\bf Projects} & {\bf OVS-eBPF } & {\bf OVS-AFXDP} \\ \hline\hline
  Packet Rate & Low & High \\ \hline
  Driver Support & No & Yes \\ \hline
  Minimal Kernel Req. & 4.18 & Unknown \\ \hline
  BPF Code Size & Large & Minimal \\ \hline
  Extensibility & Good & Great \\ \hline
  \end{tabular}
\caption{\footnotesize
Comparison of the OVS-eBPF project and OVS-AFXDP project.
}
\label{proj-compare}
\end{table}

\section{Background}
\label{sec:background}

@@

\section{OVS eBPF Datapath}
\label{sec:ebpf}
\myfigcross{ovsebpf}{5.5in}{The overall architecture of OVS eBPF datapath
consists of multiple eBPF programs which are tail-called dynamically, maps
which are shared between eBPF programs and userspace applications, and
ovs-vswitchd as the management plane for all components.}

@@

\section{Userspace OVS with AF\_XDP}
\label{sec:af_xdp}
\begin{figure}
{\scriptsize
\begin{verbatim}
              |   +-------------------+
              |   |    ovs-vswitchd   |<-->ovsdb-server
              |   +-------------------+
              |   |      ofproto      |<-->OpenFlow controllers
              |   +--------+-+--------+ 
              |   | netdev | |ofproto-|
    userspace |   +--------+ |  dpif  |
              |   | AF_XDP | +--------+
              |   | netdev | |  dpif  |
              |   +---||---+ +--------+
              |       ||     |  dpif- |
              |       ||     | netdev |
              |_      ||     +--------+  
                      ||         
               _  +---||-----+--------+
              |   | XDP program +     |
 Linux kernel |   |   eBPF map        |
              |_  +--------||---------+
                           ||
                        physical
                           NIC
\end{verbatim}
}
\vspace{-1.0em}
\caption{OVS Architecture AF\_XDP}
\label{ovsafxdp}
\vspace{-1.0em}
\end{figure}

@@

\section{Conclusion}
\label{sec:conclusion}
This paper describes two eBPF projects related to OVS: OVS-eBPF and OVS-AFXDP.
